Predicting the 2024 US Presidential Election with a Model-Based Forecast

### 2024-11-04 paper/paper.qmd
finalize paper and update scripts

The linear regression model built by our training dataset, using 242 data points, estimated the factors influencing Trump's support levels. For brevity, @tbl-summary-model only shows the first ten rows of the model’s coefficients, with the full model summary available in [Appendix -@sec-model-details]. The intercept is estimated at 44.224, representing the baseline support level when all predictors are at their reference levels. The model achieved an $R^2$ value of 0.803, indicating that 80.3% of the variance in polling outcomes is explained by the predictors included in the model. The adjusted $R^2$ value of 0.753 confirms that the model captures the relationships between variables without significant overfitting.

@@

# Load the model
model <- readRDS(file = here::here("models/first_model.rds"))

@@

Since we are missing data for some states in our dataset, we turned to historical trends from recent elections involving Joe Biden to fill these gaps. For each missing state, we assessed whether it consistently leaned toward a particular candidate in past elections. If so, we assumed this trend would persist and used it as a basis for estimating outcomes in the current election.

@@

This paper examines the factors affecting Donald Trump’s polling performance in the 2024 U.S. Presidential Election. Our linear regression model shows that numeric grade, polling methodology, and state-specific factors are key predictors of Trump’s support. Higher numeric grades, representing perceived poll reliability, are positively associated with Trump’s polling numbers, suggesting that well-regarded polls may capture a more favorable sentiment. Polling methodology also plays a significant role, with different methods (e.g., "Mixed Voting" or "Online Panel") showing varied effects on the results, likely reflecting differences in respondent engagement or accessibility. State-specific factors underscore the importance of regional dynamics. The Electoral College system makes state-level support crucial, and unique local issues and demographics in each state contribute to Trump’s polling outcomes.

@@

Beyond the mechanics of polling, our analysis suggests that state-specific policies and regional cultural factors could influence voter preferences and eventually shape electoral outcomes, although they are not directly measured in our model. For example, state policies on recreational marijuana legalization may reflect broader ideological values that align with particular candidates or parties. To explore this idea, we compared our predictions map with a visualization of states where recreational marijuana is legal.

@@

Another limitation lies in the model’s emphasis on certain predictors while excluding others. By prioritizing variables like poll duration and sample size, we may miss factors that could better explain variations in support, such as demographic and economic data. Although the model performs well on a broad level, these limitations underscore the need for more nuanced data and a more inclusive approach to fully understand the range of influences on voter behavior.

>>>>

The linear regression model built by our training dataset, using 242 data points, estimated the factors influencing Trump's support levels. For brevity, @tbl-summary-model only shows the first ten rows of the model’s coefficients, with the full model summary available in [Appendix -@sec-model-details]. The intercept is estimated at 44.224, representing the baseline support level when all predictors are at their reference levels. The model achieved an $R^2$ value of 0.803, indicating that 80.3% of the variance in polling outcomes is explained by the predictors included in the model. The adjusted $R^2$ value of 0.753 confirms that the model captures the relationships between variables without overfitting.

@@

# Load the model
model <- readRDS(file = here::here("models/linear_model.rds"))

@@

Since we are missing data for some states in our dataset, we turned to historical trends from recent elections involving Joe Biden to fill these blanks. For each missing state, we assessed whether it consistently leaned toward a particular candidate in past elections. If so, we assumed this trend would persist and used it as a basis for estimating outcomes in the current election.

@@

This paper examines the factors affecting Donald Trump’s polling performance in the 2024 U.S. Presidential Election. Our linear regression model shows that numeric grade, polling methodology, and state-specific factors are key predictors of Trump’s support. Higher numeric grades, representing perceived poll reliability, are positively associated with Trump’s polling numbers, suggesting that well-regarded polls may capture a more favorable sentiment. Polling methodology also plays an important role, with different methods (e.g., "Mixed Voting" or "Online Panel") showing varied effects on the results, likely reflecting differences in respondent engagement or accessibility. State-specific factors underscore the importance of regional dynamics. The Electoral College system makes state-level support essential, and unique local issues and demographics in each state contribute to Trump’s polling outcomes.

@@

Beyond the mechanics of polling, our analysis suggests that state-specific policies and regional cultural factors could influence voter preferences and eventually shape electoral outcomes, although they are not directly measured in our model. For example, state policies on recreational marijuana legalization may reflect broader ideological values that align with particular candidates or parties. To examine this further, we compared our predictions map with a visualization of states where recreational marijuana is legal.

@@

Another limitation lies in the model’s emphasis on certain predictors while excluding others. By prioritizing variables like poll duration and sample size, we may miss factors that could better explain variations in support, such as demographic and economic data. Although the model performs well on a broad level, these limitations underscore the need for a more inclusive approach to fully understand the range of influences on voter behavior.
